---
title: "Supervisory Resource Allocation Report"
author: "Dylan Viswambaran"
date: "27/11/2023"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE, 
                      message = FALSE)
```


```{r load_data}

source("R/main.R")

```

## Firms Overview 

To get an idea of the firms in the data. Below is a section to give a overview of the largest firms, how their profile evolves and any firms that stand out compared to its peers. 



### Gross Written Premium

Focusing on the latest year here are the top 5 firms by their Gross Written Premium (GWP):  
```{r GWP }

## Previous years average for comparison 
prevtop5Avg <-
  condensed_table_changes_long %>%
  # filter(Year == max(Year) - 1) %>% 
  group_by(Metric, Year) %>% 
  slice_max(value, n = 5) %>% 
  summarise(Average = mean(value, na.rm = TRUE)) %>% 
  ungroup()

GWPAvg <- prevtop5Avg %>% 
  filter(Metric == "GWP (£m)") %>% 
  slice_max(Year, n = 2) 

## Latest year chart and table 



latestTop5GWP <- condensed_table_changes_long %>%
  filter(Year == max(Year),
         Metric == "GWP (£m)") %>%
  slice_max(value, n = 5)



## Plot
latestTop5GWP %>%
    ggplot(aes(x = reorder(Firms, -value),
               y = value)) +
    geom_bar(stat = "identity",
             position = position_dodge2(),
             fill = boe_brand_main$boe_aqua) +
    boeCharts::theme_boe_identity_md() +
    scale_y_continuous(labels = label_currency()) +
    labs(x = "Firm",
         y = "GWP (£m)",
         title = "Top 5 Firms by Gross Written Premium") + 
    coord_flip() +
    geom_hline(yintercept = GWPAvg$Average[1], linetype = "dashed", color = "red") +
    annotate("text", x = Inf, y = GWPAvg$Average[1], label = "2020 Average (Top 5)", color = "red", hjust = -0.75, vjust = 3.5) +
    geom_hline(yintercept = GWPAvg$Average[2], linetype = "dashed", color = "yellow") +
    annotate("text", x = Inf, y = GWPAvg$Average[2], label = "2019 Average (Top 5)", color = "yellow", hjust = -0.73, vjust = 8.5)


## Table
latestTop5GWP %>% 
  select(-Metric) %>% 
  rename(`£mn` = value) %>% 
  kable(caption = "Top 5 Firms by Gross Written Premium") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"), 
                full_width = TRUE)
  

  


```
Looking at firm sizes on the average over the past few years. This will give an indication of the size of the firm over time and not just focusing on a particularly good (or bad) year. 

```{r GWPAvg}

top5AverageGWP <- condensed_table_changes_long %>%
  group_by(Firms, Metric) %>%
  summarise(Average = mean(value)) %>%
  ungroup() %>%
  filter(Metric == "GWP (£m)") %>%
  slice_max(Average, n = 5)


## Plotting average 
top5AverageGWP %>% 
  ggplot(aes(x = reorder(Firms, -Average),
             y = Average)) +
  geom_bar(stat = "identity",
           position = position_dodge2(),
           fill = boe_brand_main$boe_aqua) +
  boeCharts::theme_boe_identity_md() +
  scale_y_continuous(labels = label_currency()) +
  labs(x = "Firm",
       y = "GWP (£m)",
       title = "Top 5 Firms by Average Gross Written Premium") + 
  coord_flip()

top5AverageGWP %>% 
  select(-Metric) %>% 
  rename(`£mn` = Average) %>% 
  kable(caption = "Top 5 Firms by Average Gross Written Premium") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"), 
                full_width = TRUE)
  

```

```{r GWPChanges}
# 
# condensed_table_changes_long %>% 
#   filter(!is.infinite(value), grepl("GWP.%*|NWP|SCR|Gross claims incurred|net combined", Metric, ignore.case = TRUE)) %>% 
#   slice_max(value, n = 5)
#   
# 
# condensed_table_changes %>% 
#   select(Firms, Year, matches("GWP|NWP|SCR|Gross claims incurred|net combined")) %>% 
#   pivot_longer()


```

### Net Written Premium 



```{r NWP}

NWPAvg <- prevtop5Avg %>% 
  filter(Metric == "NWP (£m)") %>% 
  slice_max(Year, n = 2) 

## Latest year chart and table 



latestTop5NWP <- condensed_table_changes_long %>%
  filter(Year == max(Year),
         Metric == "NWP (£m)") %>%
  slice_max(value, n = 5)



## Plot
latestTop5NWP %>%
    ggplot(aes(x = reorder(Firms, -value),
               y = value)) +
    geom_bar(stat = "identity",
             position = position_dodge2(),
             fill = boe_brand_main$boe_aqua) +
    boeCharts::theme_boe_identity_md() +
    scale_y_continuous(labels = label_currency()) +
    labs(x = "Firm",
         y = "NWP (£m)",
         title = "Top 5 Firms by Net Written Premium") + 
    coord_flip() +
    geom_hline(yintercept = NWPAvg$Average[1], linetype = "dashed", color = "red") +
    annotate("text", x = Inf, y = NWPAvg$Average[1], label = "2020 Average (Top 5)", color = "red", hjust = -0.75, vjust = 3.5) +
    geom_hline(yintercept = NWPAvg$Average[2], linetype = "dashed", color = "yellow") +
    annotate("text", x = Inf, y = NWPAvg$Average[2], label = "2019 Average (Top 5)", color = "yellow", hjust = -0.73, vjust = 8.5)


## Table
latestTop5NWP %>% 
  select(-Metric) %>% 
  rename(`£mn` = value) %>% 
  kable(caption = "Top 5 Firms by Net Written Premium") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"), 
                full_width = TRUE)
  






```
### Gross claims incurred

```{r claims}

ClaimsAvg <- prevtop5Avg %>% 
  filter(Metric == "Gross claims incurred (£m)") %>% 
  slice_max(Year, n = 2) 

## Latest year chart and table 



latestTop5Claims <- condensed_table_changes_long %>%
  filter(Year == max(Year),
         Metric == "Gross claims incurred (£m)") %>%
  slice_max(value, n = 5)



## Plot
latestTop5Claims %>%
    ggplot(aes(x = reorder(Firms, -value),
               y = value)) +
    geom_bar(stat = "identity",
             position = position_dodge2(),
             fill = boe_brand_main$boe_aqua) +
    boeCharts::theme_boe_identity_md() +
    scale_y_continuous(labels = label_currency()) +
    labs(x = "Firm",
         y = "Gross claims incurred (£m)",
         title = "Top 5 Firms by Gross claims incurred") + 
    coord_flip() +
    geom_hline(yintercept = ClaimsAvg$Average[1], linetype = "dashed", color = "red") +
    annotate("text", x = Inf, y = ClaimsAvg$Average[1], label = "2020 Average (Top 5)", color = "red", hjust = -0.75, vjust = 3.5) +
    geom_hline(yintercept = ClaimsAvg$Average[2], linetype = "dashed", color = "yellow") +
    annotate("text", x = Inf, y = ClaimsAvg$Average[2], label = "2019 Average (Top 5)", color = "yellow", hjust = -0.73, vjust = 8.5)


## Table
latestTop5Claims %>% 
  select(-Metric) %>% 
  rename(`£mn` = value) %>% 
  kable(caption = "Top 5 Firms by Gross claims incurred") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"), 
                full_width = TRUE)


## Changes over time 

claimChanges <- 
  condensed_table_changes_long %>%
  filter(Metric %in% c("Gross claims incurred (£m)", "Gross claims incurred (£m) % Change"), Firms %in% c(latestTop5Claims$Firms))

claimpctChangetop5 <- 
condensed_table_changes_long %>% 
    filter(Metric == "Gross claims incurred (£m) % Change", 
           Year == max(Year), !is.infinite(value)) %>% 
    slice_max(value, n = 5)

top5pctChangeOverTime  <- 
  condensed_table_changes_long %>% 
  filter(Metric == "Gross claims incurred (£m) % Change", Firms %in% c(claimpctChangetop5$Firms), !is.na(value))


# Nominal change plot 

claimChanges %>% 
    filter(Metric == "Gross claims incurred (£m)") %>% 
    ggplot(
        aes(x = Year, 
            y = value, 
            color = Firms)
    ) + 
    geom_line() + 
  theme_boe_identity_md() +
  # theme_minimal() +
  facet_wrap(~ Firms) +
  theme(
    legend.position = "none"
  ) + 
  labs(
    y = "£mns",
    title = "Top 5 Gross claims incurred (£mn) nominal changes"
  ) + 
  scale_x_continuous(guide = guide_axis(n.dodge = 2))



# Percent change in focus chart
top5pctChangeOverTime  %>% 
    ggplot(
        aes(x = Year, 
            y = value, 
            color = Firms)
    ) + 
    geom_line() + 
  theme_boe_identity_md() + 
  facet_wrap(~ Firms, scales = "free_y") +
  theme(
    legend.position = "none"
  ) + 
  labs(
    y = "% Change",
    title = "Top 5 Gross claims incurred % changes"
  )




```




### Net combined ratio

This ratio can indicate a firms profitability. A Ratio of less than 100% can indicate that firm is in profit.
```{r NetcombinedRatio}

# Ratio to indicate profitability of a frim 


NCR <- condensed_table_changes_long %>% 
  filter(grepl("net combined ratio", Metric, ignore.case = TRUE))

NCRtoplosses <- NCR %>% 
  filter(grepl("ratio$", Metric), !is.na(value), !is.infinite(value), Year == max(Year), value > 100)
  
NCRtoplosses %>% 
    ggplot(aes(x = reorder(Firms, -value),
               y = value,
               fill = Firms)) +
    geom_bar(stat = "identity",
             position = position_dodge2()) + 
  theme_boe_identity(legend_position = "none") + 
  labs(x = "Firm",
       y = "Net Combined Ratio",
       title = "Firms with low profitability")
```
## Outlier analysis 

```{r}

# outlier_detector(condensed_table_changes_long, "value") %>%
#   pivot_wider(names_from = Metric, values_from = value) %>% 
#   kable(caption = "Table of outliers within dataset") %>% 
#   kable_styling(bootstrap_options = c("striped", "hover"), 
#                 full_width = TRUE)




condensed_table_changes_long %>%
  filter(!grepl("Change", Metric, ignore.case = TRUE)) %>% 
  ggplot(aes(
    x = Metric, 
    y = value
  )) + 
  geom_boxplot() +
  facet_wrap(~Metric, scales = "free") + 
  theme_minimal()

## Scatter plot to identify outliers visually 
condensed_table_changes_long %>%
  filter(!grepl("Change", Metric, ignore.case = TRUE)) %>% 
  ggplot(
    aes(
      x = Year,
      y = value, 
      color = Firms
    )
  ) + 
  geom_point() + 
  facet_wrap(~ Metric, scales = "free", ncol = 3, ) + 
  theme_minimal() + 
  theme(legend.position = "none")

## Focussing on latest year 
condensed_table_changes_long %>%
  filter(!grepl("Change", Metric, ignore.case = TRUE), Year == max(Year)) %>% 
  ggplot(
    aes(
      x = Firms,
      y = value, 
      color = Metric
    )
  ) + 
  geom_point() + 
  facet_wrap(~ Metric, scales = "free", ncol = 3, ) +
  theme_minimal() + 
  theme(legend.position = "none") 

```

